%!TEX encoding=UTF-8 Unicode


\documentclass{sig-alternate}

\setlength{\paperheight}{11in}
\setlength{\paperwidth}{8.5in}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{balance,tikz}

\usepackage[]{graphicx}
\graphicspath{{./img/}}

\usepackage[tight,footnotesize]{subfigure}
\usepackage[]{caption}

\usepackage[hidelinks,pdfpagelabels=false]{hyperref}

\usepackage{courier,listings}
\lstset{language=C,%
    basicstyle=\ttfamily\small,%
    captionpos=b,%
    breaklines=true,
    stepnumber=1,%
    numberstyle=\scriptsize\bfseries,
    numbersep=6pt,
    numbers=left,%
    tabsize=2,%
    escapechar=|,%
    xleftmargin=10pt,
    frame=tb,
    numbersep=2.5pt,%
}
%\captionsetup[lstlisting]{ margin=3pt }
\usetikzlibrary{decorations.pathreplacing,patterns}

\usepackage{multirow,booktabs,xspace,soul}

\newcommand{\TABARNAC}{\emph{TABARNAC}\xspace}
\newcommand{\Input}[1]{\input{./tex/#1}}

\usepackage{todonotes}
\newcommand{\tod}[1]{\textbf{\color{red}{\ul{TODO #1}}}}
\newcommand{\mytodo}[2]{\todo[inline, author=#1]{#2}}
\newcommand{\DB}[1]{\mytodo{David}{#1}}
\newcommand{\GH}[1]{\mytodo{Guillaume}{#1}}
\newcommand{\MD}[1]{\mytodo{Matthias}{#1}}

%\widowpenalty=10000
%\displaywidowpenalty=10000
%\clubpenalty=10000

\begin{document}

\title{TABARNAC: Visualizing and Resolving\\ Memory Access Issues on NUMA Architectures}

% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{4} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor David Beniamine\\
       \affaddr{Univ. Grenoble Alpes, LIG, F-38000 Grenoble, France}\\
       \affaddr{CNRS, LIG, F-38000 Grenoble, France}\\
       \affaddr{Inria}\\
       \email{David.Beniamine@Imag.fr}
% 2nd. author
\alignauthor Matthias Diener\\
    \affaddr{Informatics Institute}\\
    \affaddr{UFRGS}\\
    \affaddr{Porto Alegre, Brazil}\\
    \email{mdiener@inf.ufrgs.br}
% 3rd. author
\alignauthor
Guillaume Huard\\
       \affaddr{Univ. Grenoble Alpes, LIG, F-38000 Grenoble, France}\\
       \affaddr{CNRS, LIG, F-38000 Grenoble, France}\\
       \affaddr{Inria}\\
       \email{Guillaume.Huard@Imag.fr}
\and
%% 4th. author
\alignauthor
Philippe O. A. Navaux\\
    \affaddr{Informatics Institute}\\
    \affaddr{UFRGS}\\
    \affaddr{Porto Alegre, Brazil}\\
       \email{navaux@inf.ufrgs.br}
   }
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
%\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle


\begin{abstract}

In modern parallel architectures, memory accesses represent a common
bottleneck. This optimizing the way applications access to memory. This is even more important with NUMA
memories, as the access time to data depends on its location in
the memory. Many efforts were made to
develop adaptive tools to improve memory accesses at the runtime by optimizing
the mapping of data and threads to NUMA nodes.  However, theses tools are not
able to change the memory access pattern of the original application,
therefore a code written without considering memory performance might
not benefit from them. Moreover, automatic mapping tools take time to converge
towards the best mapping loosing optimization opportunities. A
deeper understanding of the memory behavior can help defining the best initial mapping
removing the need for runtime analysis.

In this paper, we present \TABARNAC, a tool for analyzing the memory behavior of parallel applications with a focus on NUMA architectures.
Our main contributions are to provide some visualization of the access
distribution by thread and by structure. Such visualization allows the
user to easily understand why performance issues occurs.
Using \TABARNAC, we explain why some applications do not benefit from data
and thread mapping. Moreover, we propose several simple code modifications to
improve the memory access behavior of some applications. Finally, we compare
the performance gain of our proposition to automated tools.
%
\end{abstract}
% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
%
%\terms{Theory}
%
\keywords{Memory access patterns, NUMA, Performance visualization} % NOT required for Proceedings

\Input{intro.tex}
\Input{stateOfTheArt.tex}
%\Input{motivations.tex}
\Input{design.tex}
\Input{experiments.tex}
\Input{conclusions.tex}

%\balance
\bibliography{biblio,library}
\bibliographystyle{abbrv}
\balancecolumns


\end{document}
